### 2018-06-05 Presentation/Figs/orth_pointforerec_schematic.tex
Added Presentation

\begin{tikzpicture}[x=0.7pt,y=0.7pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,505.89);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (263.88,456.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.84,106.94) --
	( 68.84,387.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 22.07,153.71) --
	(289.08,153.71);

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 68.84,153.71) -- (232.54,235.56);

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (222.59,220.48) --
	(232.54,235.56) --
	(214.51,236.64);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (238.54,233.26) {{\large ${\bm S}$}};
\definecolor{drawColor}{RGB}{0,0,255}
\definecolor{fillColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (139.00,294.02) circle (  3.00);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (139.00,312.02) {{\large $\color{blue}{\hat{\bm{y}}}$}};
\definecolor{drawColor}{RGB}{255,0,0}
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (181.09,209.84) circle (  3.00);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (181.09,186.10) {{\large $\color{red}{\tilde{\bm{y}}}$}};
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.00,294.02) -- (181.09,209.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (166.02,219.79) --
	(181.09,209.84) --
	(182.18,227.87);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (108.93,173.75) circle (  3.00);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (108.93,150.02) {{\large $\color{black}{{\bm{y}}}$}};
\end{scope}
\end{tikzpicture}