### 2018-06-05 Presentation/Figs/probforerec_schematic.tex

% Created by tikzDevice version 0.11 on 2018-04-24 21:50:28
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=0.6pt,y=0.6pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,361.35);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (480.69,312.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.68, 49.37) --
	( 88.68,302.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 91.62) --
	(505.89, 91.62);

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 88.68, 91.62) -- (417.71,165.55);

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (404.42,153.31) --
	(417.71,165.55) --
	(400.46,170.93);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (423.71,163.26) {{\Large ${\bm S}$}};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (276.70,112.53) {{\huge $\mathfrak{s}$}};

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 88.68, 91.62) -- (182.69,260.61);

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (182.98,242.54) --
	(182.69,260.61) --
	(167.19,251.33);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (182.69,266.61) {{\Large ${\bm R}$}};

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 88.68, 49.37) --
	(229.69,302.86);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (182.69, 49.37) --
	(323.70,302.86);
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 1.6pt,line join=round,line cap=round] (115.54, 97.65) --
	(222.98,121.79);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (244.09, 68.20) {{\huge $\color{red}{s(\mathcal{B})}$}};
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.69, 70.49) -- (169.26,109.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.30,108.78) --
	(169.26,109.72) --
	(177.47, 93.63);
\definecolor{fillColor}{RGB}{0,0,255}

\path[fill=fillColor,fill opacity=0.10] ( 65.18,  7.12) --
	(159.19,  7.12) --
	(347.20,345.10) --
	(253.19,345.10) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (370.70,232.76) {{\huge $\color{blue}{g^{-1}(\mathcal{B}})$}};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (370.70,218.36) -- (229.69,176.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.09,189.26) --
	(229.69,176.11) --
	(247.27,171.95);
\end{scope}
\end{tikzpicture}