### 2019-04-26 Hierarchical-forecast-reconciliation/Figs/orth_pointforerec_schematic.tex

% Created by tikzDevice version 0.12 on 2019-04-26 17:19:46
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,505.89);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (263.88,456.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.84,130.32) --
	( 68.84,317.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 22.07,177.10) --
	(209.16,177.10);

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 68.84,177.10) -- (232.54,258.95);

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (222.59,243.87) --
	(232.54,258.95) --
	(214.51,260.03);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (238.54,256.65) {{\large $\mathfrak{s}$}};
\definecolor{drawColor}{RGB}{0,0,255}
\definecolor{fillColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (139.00,317.41) circle (  3.00);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (139.00,335.41) {{\large $\color{blue}{\hat{\bm{y}}}$}};
\definecolor{drawColor}{RGB}{255,0,0}
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (181.09,233.22) circle (  3.00);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (181.09,209.48) {{\large $\color{red}{\tilde{\bm{y}}}$}};
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.00,317.41) -- (181.09,233.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (166.02,243.18) --
	(181.09,233.22) --
	(182.18,251.26);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (139.00,317.41) -- (108.93,197.14);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (103.96,214.51) --
	(108.93,197.14) --
	(121.49,210.13);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (108.93,197.14) circle (  3.00);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (108.93,173.40) {{\large $\color{black}{{\bm{y}}}$}};
\end{scope}
\end{tikzpicture}