### 2019-08-06 Application-R/Application-TourismData/Trash/Bias-correction-problem/Results.Rmd

---
title: "Bias vs Bias adjusted forecasts for Tourism data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(fpp2)
library(hts)
load("TourismData_BoxCoxTrans-BiasCorrection_UnivARIMA.RData")

```

Univariate ARIMA models were fitted for the transformed data (using BoxCox transformation). 
The back transformed data was adjusted for bias to get unbiased forecasts

Table shows the MSE for base-biased and base-unbiased forecasts for $h=1,..,6$

```{r}

DF %>% 
  group_by(`F-method`, `R-method`, Forecast_Horizon) %>% 
  summarise(MSE = mean(SquaredE)) %>% 
  ungroup() %>% 
  mutate(Forecast_Horizon = recode(Forecast_Horizon, "1" = "h=1",  "2" = "h=2",
                                    "3" = "h=3",  "4" = "h=4",  "5" = "h=5",  
                                   "6" = "h=6"),
         `F-method` = recode(`F-method`, "ARIMA_bias" = "Base-biased", 
                             "ARIMA_unbias" = "Base-unbiased")) %>% 
  select(-`R-method`) %>% 
  spread(key = Forecast_Horizon, value = MSE)

DF %>% 
  mutate(Forecast_Horizon = recode(Forecast_Horizon, "1" = "h=1",  "2" = "h=2",
                                    "3" = "h=3",  "4" = "h=4",  "5" = "h=5",  
                                   "6" = "h=6")) %>%
  group_by(`F-method`, `R-method`, Forecast_Horizon, Replication) %>% 
  summarise(MSE = mean(SquaredE)) %>% 
  spread(key = `F-method`, value = MSE) %>% 
  mutate("Bias-Unbias" = `ARIMA_bias` - `ARIMA_unbias`) %>% 
  ggplot(aes(x = Replication, y = `Bias-Unbias`)) + geom_point() + 
  facet_wrap(~Forecast_Horizon, scales = "free_y") +
  ylab("MSE (bias - unbias)")

```