### 2019-08-27 Application-R/Application-TourismData/Final-results/TourismData-Final-results.Rmd
Repeating the analysis constraining for BoxCox lambda

---
title: "Tourism Data Analysis"
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
date: "31 July 2019"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(Matrix)
library(hts)
library(readr)
library(tibble)
library(kableExtra)
# library(reshape2)
```

# Box and whisker plots for orthogonal and oblique projections

```{r, echo=FALSE}
OvernightTrips_OriginalScale_Fc <- read.csv("OvernightTrips_OriginalScale_Fc.csv")[,-1]

OvernightTrips_OriginalScale_Fc %>% 
  mutate(SquaredE = (Overnight_Trips - Overnight_Trips_Fc)^2) %>% 
  group_by(`R.method`, Fc_horizon, Replication) %>% 
  summarise(MSE = mean(SquaredE)) -> OvernightTrips_OriginalScale_MSE

DF_MSE <- DF_LogTrans %>%
  mutate(SquaredE = (`Actual` - `Forecasts`)^2) %>%
  filter(`Forecast_Horizon` == 1) %>%
  dplyr::select(-`Forecast_Horizon`) %>%
  group_by(`F.method`, `R.method`, Replication) %>%
  summarise(MSE = mean(SquaredE))

DF_MSE %>%
  filter(`F.method`=="Unbiased_M1", `R.method`%in% c("Base", "OLS", "MinT(Shrink)", "WLS")) %>%
  spread(key = `R.method`, value = MSE) %>%
  mutate("Base-OLS" = Base - OLS,
         "Base-MinT" = Base - `MinT(Shrink)`,
         "Base-WLS" = Base - WLS) %>%
  dplyr::select(Replication, `Base-OLS`, `Base-MinT`, `Base-WLS`) %>%
  rename("OLS" = "Base-OLS", "MinT" = "Base-MinT", "WLS" = "Base-WLS") %>%
  gather(`OLS`, `WLS`, `MinT`,  key = Method, value = MSE) %>%
  mutate(Method = factor(Method, levels = c("OLS", "MinT", "WLS"))) %>%
  ggplot(aes(x = Method, y = MSE)) +
  geom_boxplot() +
  ylab("MSE") + xlab("Reconciliation Method")

```

# BoxCox Transformation

```{r, echo=FALSE}
DF_BoxCoxTrans <- read.csv("DF_BoxCoxTrans.csv")[,-1]

DF_BoxCoxTrans %>% mutate(SquaredE = (`Actual` - `Forecasts`)^2) -> DF_BoxCoxTrans

DF_BoxCoxTrans %>%
  group_by(`F.method`, `R.method`, Forecast_Horizon) %>%
  summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>%
  filter(Forecast_Horizon == 1) %>%
  ungroup() %>%
  dplyr::select(-Forecast_Horizon)-> Avg_BoxCox_MSE_h1

Avg_BoxCox_MSE_h1 %>%
  mutate(F.method = recode(F.method, Unbiased_M1 = "Unbiased (Method 1)",
                           Unbiased_M2 = "Unbiased (Method 2)")) %>%
  spread(key = F.method, value = MSE) %>%
  kable(format = "latex")

```

# More analysis on BoxCox transformed results

```{r echo=FALSE}
DF_BoxCoxTrans %>% 
  group_by(`F.method`, `R.method`, Forecast_Horizon, Series) %>% 
  summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>% 
  filter(R.method == "Base", Forecast_Horizon == 1, 
         F.method %in%c("Bias","Unbiased_M1")) %>% 
  ungroup() %>% 
  dplyr::select(-R.method, -Forecast_Horizon) %>% 
  spread(key = F.method, value = MSE) %>% 
  mutate("Bias-Unbiase" = Bias - Unbiased_M1) %>% 
  filter(`Bias-Unbiase` < -100)

```

Bias correction is giving wiered results for `Total` series.

```{r echo=FALSE}
DF_BoxCoxTrans %>% 
  filter(Forecast_Horizon == 1, Series == "Total") -> DF_BoxCoxTrans_Tot_h1

# DF_BoxCoxTrans %>% 
#   filter(Forecast_Horizon == 1, Series == "A") -> DF_BoxCoxTrans_A_h1
# 
# DF_BoxCoxTrans %>% 
#   filter(Forecast_Horizon == 1, Series == "B") -> DF_BoxCoxTrans_B_h1

# DF_BoxCoxTrans_h1 %>% 
#   filter(Series == "Total", R.method == "Base", F.method == "Bias") %>% 
#   dplyr::select(Actual, Replication) -> Total_BoxCox_True

DF_BoxCoxTrans_Tot_h1 %>% 
  filter(R.method == "Base", F.method %in%c("Bias","Unbiased_M1")) %>%
  ungroup() %>% 
  dplyr::select(F.method, Forecasts, Actual, Replication) %>% 
  spread(key = F.method, value = Forecasts) %>% 
  gather(-Replication, key = "Forecasts/Actual", value = Val) -> DF_BoxCoxTrans_Tot_h1

DF_BoxCoxTrans_Tot_h1 %>% 
  mutate(`Forecasts/Actual` = recode(`Forecasts/Actual`, "Unbiased_M1" = "Unbiased")) %>%   ggplot() + 
  geom_line(aes(x = Replication, y = Val, color = `Forecasts/Actual`)) + 
  ylab("Forecasts/Actual") + 
  ggtitle("Forecasts for Total series follow from BoxCox Transformation") + 
  theme(legend.position="bottom", legend.title = element_blank())

#  DF_BoxCoxTrans_A_h1 %>% 
#   filter(R.method == "Base", F.method %in%c("Bias","Unbiased_M1")) %>%
#   ungroup() %>% 
#   dplyr::select(F.method, Forecasts, Actual, Replication) %>% 
#   spread(key = F.method, value = Forecasts) %>% 
#   gather(-Replication, key = "Forecasts/Actual", value = Val) -> DF_BoxCoxTrans_A_h1
#   
# DF_BoxCoxTrans_A_h1 %>% 
#   mutate(`Forecasts/Actual` = recode(`Forecasts/Actual`, "Unbiased_M1" = "Unbiased")) %>%   ggplot() + 
#   geom_line(aes(x = Replication, y = Val, color = `Forecasts/Actual`)) + 
#   ylab("Forecasts/Actual") + 
#   ggtitle("Forecasts for NSW series follow from BoxCox Transformation") + 
#   theme(legend.position="bottom", legend.title = element_blank()) 
#  
#  DF_BoxCoxTrans_B_h1 %>% 
#   filter(R.method == "Base", F.method %in%c("Bias","Unbiased_M1")) %>%
#   ungroup() %>% 
#   dplyr::select(F.method, Forecasts, Actual, Replication) %>% 
#   spread(key = F.method, value = Forecasts) %>% 
#   gather(-Replication, key = "Forecasts/Actual", value = Val) -> DF_BoxCoxTrans_B_h1
#   
# DF_BoxCoxTrans_B_h1 %>% 
#   mutate(`Forecasts/Actual` = recode(`Forecasts/Actual`, "Unbiased_M1" = "Unbiased")) %>%   ggplot() + 
#   geom_line(aes(x = Replication, y = Val, color = `Forecasts/Actual`)) + 
#   ylab("Forecasts/Actual") + 
#   ggtitle("Forecasts for Victoria series follow from BoxCox Transformation") + 
#   theme(legend.position="bottom", legend.title = element_blank())

```

```{r}

DF_BoxCoxTrans_Tot_h1 %>% 
  spread(key = `Forecasts/Actual`, value = Val) %>% 
  mutate("Unb-Act" = Unbiased_M1 - Actual) %>% 
  filter(`Unb-Act`>1000)

```

Very large bias corrected forecasts are given for some replications. For example in the rolling window Jul-2003 to Oct-2011. Observing the model auto.arima fits:

```{r echo=FALSE}
OvernightTrips_Region <- read_csv("OvernightTrips_2017.csv")[,-(1:3)]

## Replacing the odd value in `Adelaide Hills` with the average of its adjecent numbers ##

OvernightTrips_Region %>% 
  dplyr::select(`Adelaide Hills`) %>% 
  filter(`Adelaide Hills` > 80) %>% 
  as.numeric() -> a

which(OvernightTrips_Region$`Adelaide Hills`==a) -> x

OvernightTrips_Region %>% 
  dplyr::select(`Adelaide Hills`) %>% 
  filter(row_number() %in% c(60+12,60-12)) %>% 
  summarise(mean(`Adelaide Hills`)) %>% 
  as.numeric() -> b

OvernightTrips_Region_new <- OvernightTrips_Region %>% 
  mutate(`Adelaide Hills` = case_when(`Adelaide Hills`%% a == 0 ~ b, 
                                      TRUE ~ as.double(`Adelaide Hills`)))

#generating the hierarchy
#Hierarchy<-suppressMessages(hts(Bottom_level, list(7, c(14,21,12,12,5,5,7))))
Hierarchy <- suppressMessages(hts(OvernightTrips_Region_new, list(7, c(6,5,4,4,3,3,2), 
                                                                  c(2,2,1,4,4,1,3,1,3,6,7,3,4,3,2,3,
                                                                    3,4,2,3,1,1,1,2,2,3,4))))
AllTS_new <- allts(Hierarchy) %>%
  as_tibble()

L <- 100

```

```{r echo=FALSE}
j = 67
Train <- AllTS_new[j:(L+j-1),]
TS <- Train[,"Total"] %>% ts(frequency = 12)
fit <- auto.arima(TS, lambda = "auto", biasadj = FALSE, stepwise = FALSE, approximation = FALSE)
Fc <- forecast(fit, h=1)

Fc_trans <- Fc$mean
L_PI <- Fc$lower[,"95%"]
sig_h <- (Fc_trans - L_PI)/1.96
sig2_h <- sig_h^2

summary(fit)

```

The estimated drift term has a very large standard error. Further the variance of $\hat{y}_{t+1}$ is `r sig2_h` which is very large

# Log Transformation

```{r, echo=FALSE}
DF_LogTrans <- read.csv("DF_LogTrans.csv")[,-1]

DF_LogTrans %>% mutate(SquaredE = (`Actual` - `Forecasts`)^2) -> DF_LogTrans

DF_LogTrans %>% 
  group_by(`F.method`, `R.method`, Forecast_Horizon) %>% 
  summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>% 
  filter(Forecast_Horizon == 1) %>% 
  ungroup() %>% 
  dplyr::select(-Forecast_Horizon)-> Avg_Log_MSE_h1

Avg_Log_MSE_h1 %>% 
  mutate(F.method = recode(F.method, Unbiased_M1 = "Unbiased (Method 1)", 
                           Unbiased_M2 = "Unbiased (Method 2)")) %>% 
  spread(key = F.method, value = MSE) %>% 
  kable(format = "latex")

```