### 2019-09-02 Application-R/Application-TourismData/Final-results/TourismData-Final-results.Rmd
More on bias correction

---
title: "Tourism Data Analysis"
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
date: "31 July 2019"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(Matrix)
library(hts)
library(readr)
library(tibble)
library(kableExtra)
# library(reshape2)
```

<!-- # Box and whisker plots for orthogonal and oblique projections from better models -->

<!-- ```{r, echo=FALSE} -->
<!-- OvernightTrips_OriginalScale_Fc <- read.csv("OvernightTrips_OriginalScale_Fc.csv")[,-1] -->

<!-- OvernightTrips_OriginalScale_Fc %>% -->
<!--   mutate(SquaredE = (Overnight_Trips - Overnight_Trips_Fc)^2, -->
<!--          Fc_horizon = recode(Fc_horizon, "1" = "h=1", "2" = "h=2", -->
<!--                              "3" = "h=3", "4" = "h=4", "5" = "h=5", "6" = "h=6")) %>% -->
<!--   group_by(`R.method`, Fc_horizon, Replication) %>% -->
<!--   summarise(MSE = mean(SquaredE)) -> OvernightTrips_OriginalScale_MSE -->

<!-- OvernightTrips_OriginalScale_MSE %>% -->
<!--   filter(`R.method`%in% c("Base", "OLS", "MinT(Shrink)", "WLS")) %>% -->
<!--   spread(key = `R.method`, value = MSE) %>% -->
<!--   mutate("Base-OLS" = Base - OLS, -->
<!--          "Base-MinT" = Base - `MinT(Shrink)`, -->
<!--          "Base-WLS" = Base - WLS) %>% -->
<!--   dplyr::select(Fc_horizon, Replication, `Base-OLS`, `Base-MinT`, `Base-WLS`) %>% -->
<!--   rename("OLS" = "Base-OLS", "MinT" = "Base-MinT", "WLS" = "Base-WLS") %>% -->
<!--   gather(`OLS`, `WLS`, `MinT`,  key = Method, value = MSE) %>% -->
<!--   mutate(Method = factor(Method, levels = c("OLS", "MinT", "WLS"))) %>% -->
<!--   ggplot(aes(x = Method, y = MSE)) + -->
<!--   geom_boxplot() + facet_wrap(~ Fc_horizon) + -->
<!--   ylab("MSE") + xlab("Reconciliation Method") -->

<!-- library(scales) -->

<!-- asinh_trans <- function(){ -->
<!--   trans_new(name = 'asinh', transform = function(x) asinh(x),  -->
<!--             inverse = function(x) sinh(x)) -->
<!-- } -->

<!-- OvernightTrips_OriginalScale_MSE %>% -->
<!--   filter(`R.method`%in% c("Base", "OLS", "MinT(Shrink)", "WLS")) %>% -->
<!--   spread(key = `R.method`, value = MSE) %>% -->
<!--   mutate("Base-OLS" = Base - OLS, -->
<!--          "Base-MinT" = Base - `MinT(Shrink)`, -->
<!--          "Base-WLS" = Base - WLS) %>% -->
<!--   dplyr::select(Fc_horizon, Replication, `Base-OLS`, `Base-MinT`, `Base-WLS`) %>% -->
<!--   rename("OLS" = "Base-OLS", "MinT" = "Base-MinT", "WLS" = "Base-WLS") %>% -->
<!--   gather(`OLS`, `WLS`, `MinT`,  key = Method, value = MSE) %>% -->
<!--   mutate(Method = factor(Method, levels = c("OLS", "MinT", "WLS"))) %>% -->
<!--   ggplot(aes(x = Method, y = MSE)) + -->
<!--   geom_boxplot() + scale_y_continuous(trans = 'asinh',breaks=c(-10000,-1000,-100,0,100,1000,10000))+ -->
<!--   facet_wrap(~ Fc_horizon)+ -->
<!--   theme_bw() -->

<!-- ``` -->

<!-- # Box and whisker plots for orthogonal and oblique projections from misspecified models -->

<!-- ```{r, echo=FALSE} -->
<!-- OvernightTrips_OriginalScaleMisspecified_Fc <- read.csv("OvernightTrips_OriginalScaleMisspecified_Fc.csv")[,-1] -->

<!-- OvernightTrips_OriginalScaleMisspecified_Fc %>% -->
<!--   mutate(SquaredE = (Overnight_Trips - Overnight_Trips_Fc)^2, -->
<!--          Fc_horizon = recode(Fc_horizon, "1" = "h=1", "2" = "h=2", -->
<!--                              "3" = "h=3", "4" = "h=4", "5" = "h=5", "6" = "h=6")) %>% -->
<!--   group_by(`R.method`, Fc_horizon, Replication) %>% -->
<!--   summarise(MSE = mean(SquaredE)) -> OvernightTrips_OriginalScaleMisspecified_MSE -->

<!-- OvernightTrips_OriginalScaleMisspecified_MSE %>% -->
<!--   filter(`R.method`%in% c("Base", "OLS", "MinT(Shrink)", "WLS")) %>% -->
<!--   spread(key = `R.method`, value = MSE) %>% -->
<!--   mutate("Base-OLS" = Base - OLS, -->
<!--          "Base-MinT" = Base - `MinT(Shrink)`, -->
<!--          "Base-WLS" = Base - WLS) %>% -->
<!--   dplyr::select(Fc_horizon, Replication, `Base-OLS`, `Base-MinT`, `Base-WLS`) %>% -->
<!--   rename("OLS" = "Base-OLS", "MinT" = "Base-MinT", "WLS" = "Base-WLS") %>% -->
<!--   gather(`OLS`, `WLS`, `MinT`,  key = Method, value = MSE) %>% -->
<!--   mutate(Method = factor(Method, levels = c("OLS", "MinT", "WLS"))) %>% -->
<!--   ggplot(aes(x = Method, y = MSE)) + -->
<!--   geom_boxplot() + facet_wrap(~ Fc_horizon) + -->
<!--   ylab("MSE") + xlab("Reconciliation Method") -->

<!-- ``` -->

# Log Transformation with 100 observations of training window (Average MSE calculated over 140 replications)

```{r, echo=FALSE}
DF_LogTrans_WL100 <- read.csv("DF_LogTrans_WL100.csv")[,-1]

DF_LogTrans_WL100 %>% mutate(SquaredE = (`Actual` - `Forecasts`)^2) -> DF_LogTrans_WL100

DF_LogTrans_WL100 %>% 
  group_by(`F.method`, `R.method`, Forecast_Horizon) %>% 
  summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>% 
  filter(Forecast_Horizon == 1) %>% 
  ungroup() %>% 
  dplyr::select(-Forecast_Horizon)-> Avg_Log_WL100_MSE_h1

Avg_Log_WL100_MSE_h1 %>% 
  mutate(F.method = recode(F.method, Unbiased_M1 = "Unbiased (Method 1)", 
                           Unbiased_M2 = "Unbiased (Method 2)")) %>% 
  spread(key = F.method, value = MSE) %>% 
  kable(format = "latex")

```

# Log Transformation with 180 observations of training window (Average MSE calculated over 60 replications)

```{r, echo=FALSE}
DF_LogTrans_WL180 <- read.csv("DF_LogTrans_WL180.csv")[,-1]

DF_LogTrans_WL180 %>% mutate(SquaredE = (`Actual` - `Forecasts`)^2) -> DF_LogTrans_WL180

DF_LogTrans_WL180 %>% 
  group_by(`F.method`, `R.method`, Forecast_Horizon) %>% 
  summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>% 
  filter(Forecast_Horizon == 1) %>% 
  ungroup() %>% 
  dplyr::select(-Forecast_Horizon)-> Avg_Log_WL180_MSE_h1

Avg_Log_WL180_MSE_h1 %>% 
  mutate(F.method = recode(F.method, Unbiased_M1 = "Unbiased (Method 1)", 
                           Unbiased_M2 = "Unbiased (Method 2)")) %>% 
  spread(key = F.method, value = MSE) %>% 
  kable(format = "latex")

```

<!-- # BoxCox Transformation for $\lambda \in (0,1)$ -->

<!-- ```{r, echo=FALSE} -->
<!-- DF_BoxCoxTrans_set1 <- read.csv("OvernightTrips_BoxCoxTrans_set1_Fc.csv")[,-1] -->

<!-- DF_BoxCoxTrans_set1 %>% mutate(SquaredE = (`Actual` - `Forecasts`)^2) -> DF_BoxCoxTrans_set1 -->

<!-- DF_BoxCoxTrans_set1 %>% -->
<!--   group_by(`F.method`, `R.method`, Forecast_Horizon) %>% -->
<!--   summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>% -->
<!--   filter(Forecast_Horizon == 1) %>% -->
<!--   ungroup() %>% -->
<!--   dplyr::select(-Forecast_Horizon)-> Avg_BoxCox_set1_MSE_h1 -->

<!-- Avg_BoxCox_set1_MSE_h1 %>% -->
<!--   mutate(F.method = recode(F.method, Unbiased_M1 = "Unbiased (Method 1)", -->
<!--                            Unbiased_M2 = "Unbiased (Method 2)")) %>% -->
<!--   spread(key = F.method, value = MSE) %>% -->
<!--   kable(format = "latex") -->

<!-- ``` -->

# BoxCox Transformation for $\lambda \in (-0.5,2)$ with 100 observations of training window (Average MSE calculated over 140 replications)

```{r, echo=FALSE}
DF_BoxCoxTrans_WL100 <- read.csv("DF_BoxCoxTrans_WL100.csv")[,-1]

DF_BoxCoxTrans_WL100 %>% mutate(SquaredE = (`Actual` - `Forecasts`)^2) -> DF_BoxCoxTrans_WL100

DF_BoxCoxTrans_WL100 %>%
  group_by(`F.method`, `R.method`, Forecast_Horizon) %>%
  summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>%
  filter(Forecast_Horizon == 1) %>%
  ungroup() %>%
  dplyr::select(-Forecast_Horizon)-> Avg_BoxCox_WL100_MSE_h1

Avg_BoxCox_WL100_MSE_h1 %>%
  mutate(F.method = recode(F.method, Unbiased_M1 = "Unbiased (Method 1)",
                           Unbiased_M2 = "Unbiased (Method 2)")) %>%
  spread(key = F.method, value = MSE) %>%
  kable(format = "latex")

```

# BoxCox Transformation for $\lambda \in (-0.5,2)$ with 180 observations of training window (Average MSE calculated over 60 replications)

```{r, echo=FALSE}
DF_BoxCoxTrans_WL180 <- read.csv("DF_BoxCoxTrans_WL180.csv")[,-1]

DF_BoxCoxTrans_WL180 %>% mutate(SquaredE = (`Actual` - `Forecasts`)^2) -> DF_BoxCoxTrans_WL180

DF_BoxCoxTrans_WL180 %>%
  group_by(`F.method`, `R.method`, Forecast_Horizon) %>%
  summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>%
  filter(Forecast_Horizon == 1) %>%
  ungroup() %>%
  dplyr::select(-Forecast_Horizon)-> Avg_BoxCox_WL180_MSE_h1

Avg_BoxCox_WL180_MSE_h1 %>%
  mutate(F.method = recode(F.method, Unbiased_M1 = "Unbiased (Method 1)",
                           Unbiased_M2 = "Unbiased (Method 2)")) %>%
  spread(key = F.method, value = MSE) %>%
  kable(format = "latex")

```