### 2019-09-02 Application-R/Application-TourismData/Final-results/Trash/TourismData-Final-results.Rmd

---
title: "Tourism Data Analysis"
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
date: "31 July 2019"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(Matrix)
library(hts)
library(readr)
library(tibble)
library(kableExtra)
# library(reshape2)
```

# Box and whisker plots for orthogonal and oblique projections from better models

```{r, echo=FALSE}
OvernightTrips_OriginalScale_Fc <- read.csv("OvernightTrips_OriginalScale_Fc.csv")[,-1]

OvernightTrips_OriginalScale_Fc %>%
  mutate(SquaredE = (Overnight_Trips - Overnight_Trips_Fc)^2,
         Fc_horizon = recode(Fc_horizon, "1" = "h=1", "2" = "h=2",
                             "3" = "h=3", "4" = "h=4", "5" = "h=5", "6" = "h=6")) %>%
  group_by(`R.method`, Fc_horizon, Replication) %>%
  summarise(MSE = mean(SquaredE)) -> OvernightTrips_OriginalScale_MSE

OvernightTrips_OriginalScale_MSE %>%
  filter(`R.method`%in% c("Base", "OLS", "MinT(Shrink)", "WLS")) %>%
  spread(key = `R.method`, value = MSE) %>%
  mutate("Base-OLS" = Base - OLS,
         "Base-MinT" = Base - `MinT(Shrink)`,
         "Base-WLS" = Base - WLS) %>%
  dplyr::select(Fc_horizon, Replication, `Base-OLS`, `Base-MinT`, `Base-WLS`) %>%
  rename("OLS" = "Base-OLS", "MinT" = "Base-MinT", "WLS" = "Base-WLS") %>%
  gather(`OLS`, `WLS`, `MinT`,  key = Method, value = MSE) %>%
  mutate(Method = factor(Method, levels = c("OLS", "MinT", "WLS"))) %>%
  ggplot(aes(x = Method, y = MSE)) +
  geom_boxplot() + facet_wrap(~ Fc_horizon) +
  ylab("MSE") + xlab("Reconciliation Method")

library(scales)

asinh_trans <- function(){
  trans_new(name = 'asinh', transform = function(x) asinh(x), 
            inverse = function(x) sinh(x))
}

OvernightTrips_OriginalScale_MSE %>%
  filter(`R.method`%in% c("Base", "OLS", "MinT(Shrink)", "WLS")) %>%
  spread(key = `R.method`, value = MSE) %>%
  mutate("Base-OLS" = Base - OLS,
         "Base-MinT" = Base - `MinT(Shrink)`,
         "Base-WLS" = Base - WLS) %>%
  dplyr::select(Fc_horizon, Replication, `Base-OLS`, `Base-MinT`, `Base-WLS`) %>%
  rename("OLS" = "Base-OLS", "MinT" = "Base-MinT", "WLS" = "Base-WLS") %>%
  gather(`OLS`, `WLS`, `MinT`,  key = Method, value = MSE) %>%
  mutate(Method = factor(Method, levels = c("OLS", "MinT", "WLS"))) %>%
  ggplot(aes(x = Method, y = MSE)) +
  geom_boxplot() + scale_y_continuous(trans = 'asinh',breaks=c(-10000,-1000,-100,0,100,1000,10000))+
  facet_wrap(~ Fc_horizon)+
  theme_bw()

```

# Box and whisker plots for orthogonal and oblique projections from misspecified models

```{r, echo=FALSE}
OvernightTrips_OriginalScaleMisspecified_Fc <- read.csv("OvernightTrips_OriginalScaleMisspecified_Fc.csv")[,-1]

OvernightTrips_OriginalScaleMisspecified_Fc %>%
  mutate(SquaredE = (Overnight_Trips - Overnight_Trips_Fc)^2,
         Fc_horizon = recode(Fc_horizon, "1" = "h=1", "2" = "h=2",
                             "3" = "h=3", "4" = "h=4", "5" = "h=5", "6" = "h=6")) %>%
  group_by(`R.method`, Fc_horizon, Replication) %>%
  summarise(MSE = mean(SquaredE)) -> OvernightTrips_OriginalScaleMisspecified_MSE

OvernightTrips_OriginalScaleMisspecified_MSE %>%
  filter(`R.method`%in% c("Base", "OLS", "MinT(Shrink)", "WLS")) %>%
  spread(key = `R.method`, value = MSE) %>%
  mutate("Base-OLS" = Base - OLS,
         "Base-MinT" = Base - `MinT(Shrink)`,
         "Base-WLS" = Base - WLS) %>%
  dplyr::select(Fc_horizon, Replication, `Base-OLS`, `Base-MinT`, `Base-WLS`) %>%
  rename("OLS" = "Base-OLS", "MinT" = "Base-MinT", "WLS" = "Base-WLS") %>%
  gather(`OLS`, `WLS`, `MinT`,  key = Method, value = MSE) %>%
  mutate(Method = factor(Method, levels = c("OLS", "MinT", "WLS"))) %>%
  ggplot(aes(x = Method, y = MSE)) +
  geom_boxplot() + facet_wrap(~ Fc_horizon) +
  ylab("MSE") + xlab("Reconciliation Method")

```

# Log Transformation

```{r, echo=FALSE}
DF_LogTrans <- read.csv("OvernightTrips_LogTrans_Fc.csv")[,-1]

DF_LogTrans %>% mutate(SquaredE = (`Actual` - `Forecasts`)^2) -> DF_LogTrans

DF_LogTrans %>% 
  group_by(`F.method`, `R.method`, Forecast_Horizon) %>% 
  summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>% 
  filter(Forecast_Horizon == 1) %>% 
  ungroup() %>% 
  dplyr::select(-Forecast_Horizon)-> Avg_Log_MSE_h1

Avg_Log_MSE_h1 %>% 
  mutate(F.method = recode(F.method, Unbiased_M1 = "Unbiased (Method 1)", 
                           Unbiased_M2 = "Unbiased (Method 2)")) %>% 
  spread(key = F.method, value = MSE) %>% 
  kable(format = "latex")

```

# BoxCox Transformation for $\lambda \in (0,1)$

```{r, echo=FALSE}
DF_BoxCoxTrans_set1 <- read.csv("OvernightTrips_BoxCoxTrans_set1_Fc.csv")[,-1]

DF_BoxCoxTrans_set1 %>% mutate(SquaredE = (`Actual` - `Forecasts`)^2) -> DF_BoxCoxTrans_set1

DF_BoxCoxTrans_set1 %>%
  group_by(`F.method`, `R.method`, Forecast_Horizon) %>%
  summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>%
  filter(Forecast_Horizon == 1) %>%
  ungroup() %>%
  dplyr::select(-Forecast_Horizon)-> Avg_BoxCox_set1_MSE_h1

Avg_BoxCox_set1_MSE_h1 %>%
  mutate(F.method = recode(F.method, Unbiased_M1 = "Unbiased (Method 1)",
                           Unbiased_M2 = "Unbiased (Method 2)")) %>%
  spread(key = F.method, value = MSE) %>%
  kable(format = "latex")

```

# BoxCox Transformation for $\lambda \in (-0.5,2)$

```{r, echo=FALSE}
DF_BoxCoxTrans_set2 <- read.csv("OvernightTrips_BoxCoxTrans_set2_Fc.csv")[,-1]

DF_BoxCoxTrans_set2 %>% mutate(SquaredE = (`Actual` - `Forecasts`)^2) -> DF_BoxCoxTrans_set2

DF_BoxCoxTrans_set2 %>%
  group_by(`F.method`, `R.method`, Forecast_Horizon) %>%
  summarise(MSE = round(mean(SquaredE)/1e3, digits = 2)) %>%
  filter(Forecast_Horizon == 1) %>%
  ungroup() %>%
  dplyr::select(-Forecast_Horizon)-> Avg_BoxCox_set2_MSE_h1

Avg_BoxCox_set2_MSE_h1 %>%
  mutate(F.method = recode(F.method, Unbiased_M1 = "Unbiased (Method 1)",
                           Unbiased_M2 = "Unbiased (Method 2)")) %>%
  spread(key = F.method, value = MSE) %>%
  kable(format = "latex")

```