### 2020-01-14 R1/Letter.tex

\documentclass[11pt,a4paper]{letter}
\addtolength{\topmargin}{-2.5cm}
\addtolength{\textheight}{5cm}
\signature{George Athanasopoulos}
\date{}

\makeatletter
\def\@texttop{}
\makeatother

%\makelabels
\begin{document}
 \begin{letter}{To: Pierre Pinson\\
 	Editor, International Journal of Forecasting\\
 	DTU Elektro\\
 	348, {\O}rsteds Pl., \\
 	2800 Kongens Lyngby, Denmark}
    From: George Athanasopoulos\\
    Department of Econometrics and Business Statistics\\
    Monash University\\
    Melbourne, VIC, 3800, Australia

 \opening{ March 12, 2018\\ \\Dear Professor Pinson,}
%\textbf{RE: Revision for Computational and Graphical Statistics by Tan Ban Kheng, Anastasios
%	Panagiotelis and George Athanasopoulos}
 \medskip

Please find attached a revised manuscript for ``Forecast reconciliation: a geometric view with new insights on bias correction" by Anastasios Panagiotelis, Puwasala Gamakumara, George Athanasopoulos and Rob Hyndman for your consideration for publication in International Journal of Forecasting. We are very excited to have been given the opportunity to revise and resubmit. We carefully considered the comments offered by the two reviewers. Herein, we briefly summarize the major changes in the revision:
\begin{enumerate}
	\item blah
\end{enumerate}

The comments made by the referees warrant a detailed response which is appended to this letter. Overall, we believe we have now addressed all the concerns raised. We thank the referees for their comments as addressing these has significantly improved our manuscript. We look forward to hearing from you.

 \closing{With kind regards and best wishes,}

 \end{letter}
\end{document}