### 2020-01-17 Application-R/Application-TourismData/Final-results/TourismData-Final-resultsR1.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Tourism Data Analysis},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Tourism Data Analysis}
\author{}
\date{\vspace{-2.5em}Jan 2020}

\begin{document}
\maketitle

\hypertarget{log-transformation-with-100-observations-of-training-window-me-and-mse-calculated-over-140-replications}{%
\section{Log Transformation with 100 observations of training window (ME
and MSE calculated over 140
replications)}\label{log-transformation-with-100-observations-of-training-window-me-and-mse-calculated-over-140-replications}}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\
\hline
Base & 99.98 & 52.70 & 140.46\\
\hline
Bottom-up & 176.31 & 64.42 & 271.73\\
\hline
MinT(Shrink) & 74.31 & 50.50 & 109.89\\
\hline
OLS & 50.73 & 38.64 & 65.82\\
\hline
WLS & 121.11 & 57.50 & 174.35\\
\hline
\end{tabular}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\
\hline
Base & 4.47 & 4.43 & 4.51\\
\hline
Bottom-up & 6.36 & 5.26 & 8.06\\
\hline
MinT(Shrink) & 4.32 & 4.16 & 4.61\\
\hline
OLS & 4.34 & 4.31 & 4.36\\
\hline
WLS & 4.82 & 4.38 & 5.43\\
\hline
\end{tabular}

\hypertarget{log-transformation-with-180-observations-of-training-window-me-and-mse-calculated-over-60-replications}{%
\section{Log Transformation with 180 observations of training window (ME
and MSE calculated over 60
replications)}\label{log-transformation-with-180-observations-of-training-window-me-and-mse-calculated-over-60-replications}}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\
\hline
Base & 90.45 & 67.34 & 109.95\\
\hline
Bottom-up & 152.69 & 98.06 & 197.11\\
\hline
MinT(Shrink) & 64.64 & 59.36 & 85.54\\
\hline
OLS & 41.62 & 35.92 & 47.51\\
\hline
WLS & 110.65 & 79.64 & 136.58\\
\hline
\end{tabular}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\
\hline
Base & 4.89 & 4.78 & 5.08\\
\hline
Bottom-up & 9.71 & 6.97 & 12.70\\
\hline
MinT(Shrink) & 4.75 & 4.63 & 5.35\\
\hline
OLS & 4.67 & 4.60 & 4.81\\
\hline
WLS & 6.33 & 5.22 & 7.58\\
\hline
\end{tabular}

\hypertarget{boxcox-transformation-for-lambda-in--0.52-with-100-observations-of-training-window-me-and-mse-calculated-over-140-replications}{%
\section{\texorpdfstring{BoxCox Transformation for
\(\lambda \in (-0.5,2)\) with 100 observations of training window (ME
and MSE calculated over 140
replications)}{BoxCox Transformation for \textbackslash lambda \textbackslash in (-0.5,2) with 100 observations of training window (ME and MSE calculated over 140 replications)}}\label{boxcox-transformation-for-lambda-in--0.52-with-100-observations-of-training-window-me-and-mse-calculated-over-140-replications}}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\
\hline
Base & 94.38 & 59.71 & 129.50\\
\hline
Bottom-up & 163.07 & 71.22 & 242.39\\
\hline
MinT(Shrink) & 71.10 & 57.94 & 96.92\\
\hline
OLS & 50.70 & 43.13 & 70.16\\
\hline
WLS & 113.67 & 65.49 & 157.17\\
\hline
\end{tabular}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\
\hline
Base & 4.59 & 4.52 & 4.78\\
\hline
Bottom-up & 6.21 & 5.31 & 7.46\\
\hline
MinT(Shrink) & 4.29 & 4.17 & 4.54\\
\hline
OLS & 4.46 & 4.40 & 4.63\\
\hline
WLS & 4.77 & 4.42 & 5.23\\
\hline
\end{tabular}

\hypertarget{boxcox-transformation-for-lambda-in--0.52-with-180-observations-of-training-window-me-and-mse-calculated-over-60-replications}{%
\section{\texorpdfstring{BoxCox Transformation for
\(\lambda \in (-0.5,2)\) with 180 observations of training window (ME
and MSE calculated over 60
replications)}{BoxCox Transformation for \textbackslash lambda \textbackslash in (-0.5,2) with 180 observations of training window (ME and MSE calculated over 60 replications)}}\label{boxcox-transformation-for-lambda-in--0.52-with-180-observations-of-training-window-me-and-mse-calculated-over-60-replications}}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\
\hline
Base & 88.91 & 73.25 & 103.24\\
\hline
Bottom-up & 142.37 & 98.60 & 178.09\\
\hline
MinT(Shrink) & 63.30 & 62.84 & 75.61\\
\hline
OLS & 47.03 & 44.71 & 51.25\\
\hline
WLS & 105.78 & 83.93 & 124.99\\
\hline
\end{tabular}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\
\hline
Base & 4.61 & 4.58 & 4.72\\
\hline
Bottom-up & 9.17 & 7.00 & 11.43\\
\hline
MinT(Shrink) & 4.50 & 4.53 & 4.82\\
\hline
OLS & 4.43 & 4.43 & 4.52\\
\hline
WLS & 6.24 & 5.45 & 7.11\\
\hline
\end{tabular}

\end{document}