### 2020-01-21 Application-R/Application-TourismData/Final-results/TourismData-Final-resultsR1.tex
Adding New Table

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Tourism Data Analysis},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Tourism Data Analysis}
\author{}
\date{\vspace{-2.5em}Jan 2020}

\begin{document}
\maketitle

\hypertarget{log-transformation-with-100-observations-of-training-window-me-and-mse-calculated-over-140-replications}{%
\section{Log Transformation with 100 observations of training window (ME
and MSE calculated over 140
replications)}\label{log-transformation-with-100-observations-of-training-window-me-and-mse-calculated-over-140-replications}}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\\hline
Base & 1.00 & 0.53 & 1.40\\ \hline
Bottom-up & 1.76 & 0.64 & 2.72\\ \hline
MinT(Shrink) & 0.74 & 0.51 & 1.10\\\hline
OLS & 0.51 & 0.39 & 0.66\\\hline
WLS & 1.21 & 0.58 & 1.74\\
\hline
\end{tabular}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\\hline
Base & 1.00 & 0.99 & 1.01\\\hline
Bottom-up & 1.42 & 1.18 & 1.80\\\hline
MinT(Shrink) & 0.97 & 0.93 & 1.03\\\hline
OLS & 0.97 & 0.96 & 0.98\\\hline
WLS & 1.08 & 0.98 & 1.21\\\hline
\end{tabular}

\hypertarget{boxcox-transformation-for-lambda-in--0.52-with-100-observations-of-training-window-me-and-mse-calculated-over-140-replications}{%
\section{\texorpdfstring{BoxCox Transformation for
\(\lambda \in (-0.5,2)\) with 100 observations of training window (ME
and MSE calculated over 140
replications)}{BoxCox Transformation for \textbackslash lambda \textbackslash in (-0.5,2) with 100 observations of training window (ME and MSE calculated over 140 replications)}}\label{boxcox-transformation-for-lambda-in--0.52-with-100-observations-of-training-window-me-and-mse-calculated-over-140-replications}}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\\hline
Base & 1.00 & 0.63 & 1.37\\\hline
Bottom-up & 1.73 & 0.75 & 2.57\\\hline
MinT(Shrink) & 0.75 & 0.61 & 1.03\\\hline
OLS & 0.54 & 0.46 & 0.74\\\hline
WLS & 1.20 & 0.69 & 1.67\\\hline
\end{tabular}

\begin{tabular}{l|r|r|r}
\hline
R.method & Biased & Unbiased (Method 1) & Unbiased (Method 2)\\
\hline
Base & 1.00 & 0.98 & 1.04\\\hline
Bottom-up & 1.35 & 1.16 & 1.63\\\hline
MinT(Shrink) & 0.93 & 0.91 & 0.99\\\hline
OLS & 0.97 & 0.96 & 1.01\\\hline
WLS & 1.04 & 0.96 & 1.14\\\hline
\end{tabular}

\end{document}