### 2020-01-29 Hierarchical-forecast-reconciliation/Tables/lossvmethods_h_3.tex
moved filed around

% latex table generated in R 3.6.2 by xtable 1.8-4 package
% Thu Jan 30 00:24:36 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \hline
Loss Function & Base & Bottom-up & OLS & Structural-WLS & Spend-WLS & MinT \\ 
  \hline
TSE & 1.00 & 1.25 & 0.97 & 1.16 & 0.98 & 0.97 \\ 
  Structural WSE & 1.00 & 1.00 & 0.96 & 0.98 & 1.00 & 0.93 \\ 
  Spend WSE & 1.00 & 1.23 & 0.97 & 1.15 & 0.98 & 0.96 \\ 
   \hline
\end{tabular}
\caption{Means of different loss functions for 3-step ahead forecasts using different reconciliation methods in the tourism application.  All figures are reported relative to base forecasts.} 
\end{table}